\noindent\rule{13cm}{1.0pt}\\ \vspace{0.2cm} A Polynomial-time Reduction of the 3-SAT to the Quadratic Congruence Problem and other Related Problems \noindent\rule{13cm}{0.8pt}

### 2017-11-17 paper/paper.tex
Reduction complete.

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{thmtools,thm-restate}
\usepackage{amsfonts}
\usepackage{hyperref}
\usepackage[backend=biber,url=true,doi=true,eprint=false,style=alphabetic]{biblatex}
\usepackage{enumitem}
\usepackage[justification=centering,singlelinecheck=false]{caption}
\usepackage{indentfirst}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage[x11names, rgb]{xcolor}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{linegoal}
\usepackage{csquotes}
\usetikzlibrary{snakes,arrows,shapes}

@@

\newcounter{dummy-def}\numberwithin{dummy-def}{section}
\newtheorem{definition}[dummy-def]{Definition}
\newtheorem{theorem}{Theorem}
\newcounter{dummy-prop}\numberwithin{dummy-prop}{section}
\newtheorem{proposition}[dummy-prop]{Proposition}
\newcounter{dummy-corollary}\numberwithin{dummy-corollary}{section}
\newtheorem{corollary}[dummy-corollary]{Corollary}
\newcounter{dummy-lemma}\numberwithin{dummy-lemma}{section}
\newtheorem{lemma}[dummy-lemma]{Lemma}
\newcounter{dummy-ex}\numberwithin{dummy-ex}{section}
\newtheorem{exercise}[dummy-ex]{Exercise}
\newcounter{dummy-eg}\numberwithin{dummy-eg}{section}
\newtheorem{example}[dummy-eg]{Example}

\numberwithin{equation}{section}

@@

From this we know that the 3-SAT is NP-complete. Not only that, any problem that is reduces
3-SAT is also in NP\@. In the next sections we will provide a brief review on polynomial-time
reductions, present proper definitions on the QCP and 3-SAT, and prove the reduction. This section
provided a brief historical overview, but we do not restrict historical remarks to only this
section. When relevant, we provide short historical anecdotes on the subject.

@@

>>>>

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{thmtools,thm-restate}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage[backend=biber,url=true,doi=true,eprint=false,style=alphabetic]{biblatex}
\usepackage{enumitem}
\usepackage[justification=centering,singlelinecheck=false]{caption}
\usepackage{indentfirst}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage[x11names, rgb]{xcolor}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{linegoal}
\usepackage{csquotes}
\usetikzlibrary{snakes,arrows,shapes}

@@

\newcounter{dummy-def}\numberwithin{dummy-def}{section}
\newtheorem{definition}[dummy-def]{Definition}
\newtheorem{theorem}{Theorem}
\newcounter{dummy-prop}\numberwithin{dummy-prop}{section}
\newtheorem{proposition}[dummy-prop]{Proposition}
\newcounter{dummy-corollary}\numberwithin{dummy-corollary}{section}
\newtheorem{corollary}[dummy-corollary]{Corollary}
\newtheorem{lemma}{Lemma}
\newcounter{dummy-ex}\numberwithin{dummy-ex}{section}
\newtheorem{exercise}[dummy-ex]{Exercise}
\newcounter{dummy-eg}\numberwithin{dummy-eg}{section}
\newtheorem{example}[dummy-eg]{Example}

@@

From this we know that the 3-SAT is NP-complete. Not only that, any problem that reduces 3-SAT is
also in NP\@. In the next sections we will provide a brief review on polynomial-time reductions,
present proper definitions on the QCP and 3-SAT, and prove the reduction. This section provided a
brief historical overview, but we do not restrict historical remarks to only this section. When
relevant, we provide short historical anecdotes on the subject.

@@

In order for $\phi_R$ to be satisfiable, every $\sigma_k\in\phi_R$ must be satisfiable. This is an
if and only if condition on $y_k\in\{0,1,2,3\}$, as we showed earlier. For any satisfiable
$\phi_R$, we have the following inequalities.

\begin{align*}
  -3&\leq R_k\leq 4, \quad &k=1,2,\ldots,m\\
  0&\leq R_0\leq 2
\end{align*}

From this, it follows that

\begin{equation}\label{eq-1}
  R_k=0, \quad k=0,1,\ldots,m \iff \sum_{k=0}^m R_k\cdot 8^k = 0
\end{equation}

and also that

\begin{equation}\label{eq-2}
  \left| \sum_{k=0}^m R_k\cdot 8^k \right| < 8^{m+1}
\end{equation}

This is true because

\begin{align*}
  \left|\sum_{k=0}^m R_k\cdot 8^k\right|=\left|R_0\cdot 8^0+\sum_{k=1}^m R_k\cdot 8^k\right|&\leq
    2+\sum_{k=1}^m 4\cdot 8^k\\
    &\leq 2+(4\cdot 8+4\cdot 8^2+\cdots+4\cdot 8^m)\\
    &\leq2+8\underbrace{(4+4\cdot 8+4\cdot 8^2+\cdots+4\cdot 8^{m-1})}_{\text{geometric series}}\\
    &\leq 2+ 8\left(\frac{4}{7}\cdot 8^m-\frac{4}{7}\right)\\
    &\leq 2+\frac{4}{7}\cdot 8^{m+1} -8\cdot\frac{4}{7}\\
    &< 8^{m+1}
\end{align*}

From (\ref{eq-1}) and (\ref{eq-2}), we have that

\begin{equation}\label{eq-3}
  R_k=0,\quad k=0,1,\ldots,m \iff \sum_{k=0}^m R_k\cdot 8^k\equiv 0\mod 8^{m+1}.
\end{equation}

We now wish to prove that we can reorganize the RHS of~\autoref{eq-3} into the following expression
by using the terms from the algorithm. That is, that

\begin{equation}\label{eq-p2}
 \sum_{k=0}^m R_k\cdot 8^k\equiv 0\mod 8^{m+1}\iff\sum_{j=0}^n c_j\alpha_j\equiv\tau\mod 8^{m+1}.
\end{equation}

\begin{proof}[Proof of \autoref{eq-p2}]

We replace variables $y_k$ and valuations $r(x_i)$ by $\{-1,+1\}$-valued variables:

\begin{align*}
  y_k=&\frac{1}{2}\left[\left(1-\alpha_{2k-1}\right)+2\cdot\left(1-\alpha_{2k}\right)\right]\\
  r(x_i)=&\frac{1}{2}\left(1-\alpha_{2m+i}\right)
\end{align*}

We call $R_k'$ the result of this substitution. It follows that

\begin{equation*}
  R_k'=\frac{1}{2}[(1-\alpha_{2k-1})+2(1-\alpha_{2k})]-\sum_{x_i\in\sigma_k}\frac{1}{2}(1-
  \alpha_{2m+i})-\sum_{\ov{x}_i\in\sigma_k}\left[1-\frac{1}{2}(1-\alpha_{2m+i})\right]+1
\end{equation*}

when $\sigma_k\in\phi_R$. Plugging $R_k'$ into (\ref{eq-3}) gives us

\begin{equation}\label{eq-4}
  \sum_{k=0}^m\left(\frac{8^k}{2}-\frac{8^k}{2}\alpha_{2k-1}+8^k-8^k\alpha_{2k}-
  \sum_{x_i\in\sigma_k}\left(\frac{8^k}{2}-\frac{8^k}{2}\alpha_{2m+i}\right)-
  \sum_{\ov{x}_i\in\sigma_k}\left(\frac{8^k}{2}+\frac{8^k}{2}\alpha_{2m+i}\right)+8^k\right)\equiv
  0\mod 8^{m+1}
\end{equation}

Recall the $c_j$, $j=0,1,\ldots,n$ variables computed earlier in the algorithm:

\begin{align*}
  &\begin{rcases*}
    c_{2k-1}=-\frac{1}{2}8^k\\
    c_{2k}=-8^k\\
  \end{rcases*} &j=1,\ldots,2m\\
  &\;c_{2m+j}=\frac{1}{2}(f_j^+-f_j^-), &j=1,2,\ldots,l
\end{align*}

We wish to take equation (\ref{eq-4}) and make it look like

\begin{equation*}
  \sum_{j=0}^n c_j\alpha_j\equiv\tau\mod 8^{m+1}, \quad\alpha_j\in\{-1,+1\}.
\end{equation*}

Replacing the values from (\ref{eq-4}) with the corresponding $c_j$ values gives us

\begin{equation}\label{eq-5}
  \sum_{k=0}^m\left(\underbrace{-c_{2k-1}+c_{2k-1}\alpha_{2k-1}-c_{2k}+c_{2k}\alpha_{2k}}_{(\star)}
  -\underbrace{\sum_{x_i\in\sigma_k}8^k\left(\frac{1}{2}-\frac{1}{2}\alpha_{2m+i}\right)-\sum_{
  \ov{x}_i\in\sigma_k}8^k\left(\frac{1}{2}+\frac{1}{2}\alpha_{2m+i}\right)}_{(\star\star)}+
  \underbrace{8^k}_{(\star\star\star)}\right)
\end{equation}

on the LHS\@. We first turn our attention to $(\star)$. Notice how the pair $(2k-1,2k)$ over the
range $[0,m]$ covers the range $[0,2m]$, meaning that

\begin{equation*}
  \sum_{k=0}^m \left(c_{2k-1}+c_{2k}\right) = \sum_{i=0}^{2m} c_{i}.
\end{equation*}

In particular, separating the terms of the summation into two parts gives us

\begin{equation*}
  \overbrace{\sum_{k=0}^m\left(c_{2k-1}\alpha_{2k-1}\right)-\sum_{k=0}^m\left(c_{2k-1}+c_{2k}
  \right)}^{(\star)} = \sum_{j=0}^{2m}c_j\alpha_j-\sum_{j=0}^{2m}c_j
\end{equation*}

Recall from the algorithm that

Since $f_j^+$ and $f_j^-$ are both counting the number of literals in $\phi_R$, it is easy to see
that $\sum_{x_i\in\phi_R}f_i^+=\sum_{\ov{x}_i\in\phi_R}f_i^-$, since they range over all
$\sigma_k$, and $\Sigma$ is the set of all $\sigma_k$ relevant clause permutations. So every
literal $x_i$ must appear the same number of times as $\ov{x}_i$. This means $\sum_{j=1}^l c_{2m+j}
= 0$. Note how there exactly $l$ variables in $\phi_R$. From this we have that

\begin{equation*}
  \sum_{j=0}^{2m}c_j\alpha_j-\sum_{j=0}^{2m}c_j = \sum_{j=0}^n c_j\alpha_j-\sum_{j=0}^n c_j.
\end{equation*}

We now consider $(\star\star)$. It follows from $(\star\star)$ that

\begin{equation*}
  -\underbrace{\sum_{x_i\in\sigma_k}\frac{8^k}{2}}_{\frac{f_i^+}{2}}-
  \underbrace{\sum_{\ov{x}_i\in\sigma_k}\frac{8^k}{2}}_{\frac{f_i^-}{2}}-\frac{8^k}{2}
  \underbrace{\left(\sum_{x_i\in\sigma_k}\alpha_{2m+i}-\sum_{\ov{x}_i\in\sigma_k}\alpha_{2m+i}
  \right)}_{=0\text{ when summing over $[0,m]$}} = -\sum_{\hat{x}_i\in\sigma_k}
  \frac{1}{2}\left(f_i^++f_i^-\right)
\end{equation*}

where we say $\hat{x}_i\in\sigma_k$ if $x_i\in\sigma_k$ or $\ov{x}_i\in\sigma_k$. The same argument
from $(\star)$ can be used to show that

\begin{equation*}
  -\sum_{k=0}^m\sum_{\hat{x}_i\in\sigma_k}\frac{1}{2}(f_i^++f_i^-)=-\sum_{k=0}^m\sum_{\hat{x}_i
  \in\sigma_k}f_i^-
\end{equation*}

Now we have a summation over all disjunctive clauses $\sigma_k$, and for each $k$ we are summing
all occurrences in which $\hat{x}_i\in\sigma_k$. But this means we are summing all times in which
each variable occurs in all clauses. The number of variables is $l$, so it follows that

\begin{equation*}
  -\sum_{k=0}^m\sum_{\hat{x}_i\in\sigma_k}f_i^- =-\sum_{i=1}^l f_i^-.
\end{equation*}

Let us now focus on $(\star\star\star)$. It is easy to see that

\begin{equation*}
  \sum_{k=0}^m 8^k = \sum_{\sigma_j\in\phi_R} 8^j = -\tau_\phi
\end{equation*}

Plugging the results (note that these results already take into account the outer summation
$\sum_{k=0}^m$, and thus we ommit it) from $(\star), (\star\star), (\star\star\star)$ in
\autoref{eq-5} gives us the following expression:

\begin{equation*}
  \underbrace{\sum_{j=0}^n c_j\alpha_j-\sum_{j=0}^n c_j}_{(\star)}-\underbrace{\sum_{i=1}^l
  f_i^-}_{(\star\star)}-\underbrace{\tau_\phi}_{(\star\star\star)}\equiv 0\mod 8^{m+1}
\end{equation*}

Isolating the first summation of $(\star)$ yields

\begin{equation*}
  \sum_{j=0}^n c_j\alpha_j\equiv\tau_\phi+\sum_{j=0}^n c_j+\sum_{i=1}^l f_i^-\mod 8^{m+1}
\end{equation*}

The RHS of the congruence is exactly the definition of $\tau$ from the algorithm. Therefore:

\begin{equation*}
  \sum_{j=0}^n c_j\alpha_j\equiv\tau\mod 8^{m+1}, \quad\alpha_j\in\{-1,+1\}
\end{equation*}

\end{proof}

From the definition of $\theta_j$, $j=0,\ldots,n$, it follows directly that the above is equivalent
to

\begin{equation*}
  \sum_{j=0}^n\theta_j\alpha_j\equiv\tau\mod 8^{m+1}, \quad\alpha_j\in\{-1,+1\}
\end{equation*}

\begin{lemma}\label{lemma-1}
  Let $K$ and $H$ be as in the algorithm. The general solution of the system
  \begin{align*}
    &0\leq|x|\leq H, \qquad x\in\mathbb{Z} \quad &(1)\\
    &(H+x)(H-x)\equiv 0\mod K &(2)
  \end{align*}
  is given by
  \begin{equation*}
    x=\sum_{j=0}^n\alpha_j\theta_j, \quad\alpha_j\in\{-1,+1\},\quad j=0,1,\ldots,n.
  \end{equation*}
\end{lemma}
\begin{proof}
  Let us recall the definitions of $K$ and $H$ from the algorithm:

  \begin{align*}
    &K=\prod_{i=0}^n p_i^{n+1} \quad\text{, where $p_0=13$ and $p_1,p_2,\ldots,p_n$ primes
    exceeding 12.}\\
    &H=\sum_{i=0}^n\theta_i
  \end{align*}

  It is obvious that $x$ is a solution to the system. We need to prove that these are the only
  solutions to the system. Let $x$ be a solution to the system. Then

  \begin{equation*}
    (H+x)(H-x)\equiv 0\mod p_j^{n+1}, \qquad j=0,1,\ldots,n
  \end{equation*}

  We will prove by contradiction. Assume that for some $j_0$:

  \begin{equation*}
    p_{j_0}|(H+x)\text{ and }p_{j_0}|(H-x)
  \end{equation*}

  Then $p_{j_0}|(H+x)+(H-x)=2H$. But $p_{j_0}>2$ and $p_{j_0}$ is prime, so $p_{j_0}|H$. It then
  follows that $p_{j_0}|\sum_{j=0}^n\theta_j$. But from the definition of $\theta_j$, we know that
  $p_{j_0}|\theta_j$ for all $j\neq j_0$, since we have

  \begin{equation*}
    \theta_j\equiv 0\mod\prod_{i\neq j}^n p_i^{n+1}\implies\prod_{i\neq j}^n p_i^{n+1}|\theta_j
    \implies p_i|\theta_j, \quad i\neq j
  \end{equation*}

  So it must be that $p_{j_0}|\theta_{j_0}$, but from $\theta_j$'s definition:

  \begin{equation*}
    \theta_j\not\equiv 0\mod p_j\implies p_j\nmid\theta_j
  \end{equation*}

  We end up in a contradiction. Therefore we can conclude that for all $j$, $p_j^{n+1}$ divides
  either $(H+x)$ or $(H-x)$, but not both.

  We define:

  \begin{align*}
    &\alpha_j=
    \begin{cases}
      \phantom{-}1,  &\quad \text{if } p_j^{n+1}|(H-x)\\
      -1,  &\quad \text{if } p_j^{n+1}|(H+x)
    \end{cases}\\
    &x'=\sum_{j=0}^n\alpha_j\theta_j
  \end{align*}

  So we have $x'\equiv\alpha_j\theta_j\equiv\alpha_j H\equiv x\mod p_j^{n+1}$, and $x'\equiv x\mod
  p_j^{n+1}$, for all $j$, and so $x'\equiv x\mod K$.

  \begin{equation*}
    \begin{rcases}
      & -H\leq x'\leq H\\
      & -H\leq x\leq H\\
    \end{rcases} \implies |x-x'|\leq 2H
  \end{equation*}

  But $p_j\geq{(4(n+1)8^{m+1})}^\frac{1}{n+1}$, and let

  \begin{equation*}
    \lambda_j=\frac{\theta_j}{\sum_{\substack{i=0\\i\neq j}}^n p_i^{n+1}}
  \end{equation*}

  We have that $\lambda_j<2\cdot 8^{m+1}$ for each $j$. Each term in $H$ is bounded by $K/2(n+1)$,
  so $2H<K$. This means $|x-x'|<K$, but $x'\equiv x\mod K$. Therefore, it must be that $x=x'$.

  Thus $x'$ is a solution and is the same as $x$, meaning any solution of the system is of that
  form.
\end{proof}

From~\autoref{lemma-1}, the condition

\begin{equation*}
  \sum_{j=0}^n\theta_j\alpha_j\equiv\tau\mod 8^{m+1}
\end{equation*}

is equivalent to the system:

\begin{align*}
  (i) \quad &0\leq|x|\leq H, \quad x\in\mathbb{Z},\\
  (ii) \quad &x\equiv\tau\mod 8^{m+1}, &\text{(I)}\\
  (iii) \quad &(H+x)(H-x)\equiv 0\mod K
\end{align*}

\begin{lemma}\label{lemma-2}
  Let $\tau$ be odd, $x\in\mathbb{Z}$, $k\geq 3$
  \begin{align*}
    (\tau+x)(\tau-x)\equiv 0\mod 2^{k+1}\iff\text{ either }\;&\tau+x\equiv 0\mod 2^k\\
                                            \text{ or }\;&\tau-x\equiv 0\mod 2^k
  \end{align*}
\end{lemma}
\begin{proof}
  We first prove the converse $(\impliedby)$:

  Assume it is possible for both cases to be true.
  \begin{align*}
    &\tau+x\equiv 0\mod 2^k\implies\tau\equiv -x\mod 2^k\\
    &\tau-x\equiv 0\mod 2^k\implies\tau\equiv \phantom{-}x\mod 2^k
  \end{align*}
  So $\tau\equiv x\equiv 0\mod 2^k$. But $\tau$ is odd, leading to a contradiction. Therefore, only
  one of the two can be true at the same time. Consider the first case ($\tau+x\equiv 0\mod 2^k$).
  So $(\tau+x)\equiv 0\mod 2^k\implies(\tau+x)(\tau-x)\equiv 0\mod 2^k$. In other words,
  $(\tau+x)(\tau-x)= 2^k \cdot c$. Multiplying $2$ on both sides yields $2(\tau+x)(\tau-x)
  =2^{k+1}\cdot c$. Converting it back to its congruence form gives us $2(\tau+x)(\tau-x)\equiv
  0\mod 2^{k+1}$. But $\tau+x\equiv 0\mod 2^k$, so we can substitute $(\tau+x)$ for $2^k$, giving
  $2\cdot 2^k\cdot c(\tau-x)=2^{k+1}c(\tau-x)\equiv 0\mod 2^{k+ 1}$.  We do the same for
  $\tau-x\equiv 0\mod 2^k$.

  We now prove the direct implication $(\implies)$:

  \begin{align*}
    (\tau+x)(\tau-x)\equiv 0\mod 2^{k+1}\implies\text{ either }\;&\tau+x\equiv 0\mod 2^k\\
                                            \text{ or }\;&\tau-x\equiv 0\mod 2^k
  \end{align*}

  We can rewrite $(\tau+x)(\tau-x)\equiv 0\mod 2^{k+1}$ as $(\tau+x)(\tau-x)=2^{k+1}\cdot c=2^k(2c)
  =(\tau+x)(\tau-x)$. So $(\tau+x)(\tau-x)\equiv 0\mod 2^k$.

  Let $a=\tau+x$ and $b=\tau-x$. We need to show that the cases:
  \begin{align*}
    (1) \quad &a\equiv 0\mod 2^k\quad\text{and}\quad b\neq 0\mod 2^k\\
    (2) \quad &b\not\equiv 0\mod 2^k\quad\text{and}\quad b\not\equiv 0\mod 2^k
  \end{align*}
  are impossible if $a\cdot b\equiv 0\mod 2^k$. We take case (1) into consideration first.

  \begin{equation*}
    \begin{cases}
      &\tau+x\equiv 0\mod 2^k\\
      &\tau-x\equiv 0\mod 2^k\\
    \end{cases}
  \end{equation*}

  But from the converse's proof, we know that this is impossible. Now we consider the second case.

  \begin{align*}
    \begin{cases}
      &\tau+x\not\equiv 0\mod 2^k\\
      &\tau-x\not\equiv 0\mod 2^k\\
      &(\tau+x)(\tau-x)\equiv 0\mod 2^k\\
    \end{cases}\quad&\implies&
    \begin{cases}
      &2^k\nmid\tau+x\\
      &2^k\nmid\tau-x\\
      &2^k\nmid(\tau+x)(\tau-x)\\
    \end{cases}
  \end{align*}

  Which is a contradiction. Therefore, the two cases are impossible. It is easy to see that all the
  remaining cases are the ones listed in the Lemma, and all are true.
\end{proof}

Consider the system (I). The conditions of~\autoref{lemma-2} apply to the system. Therefore system
(I) is satisfiable if and only if system (II) is also satisfiable:

\begin{align*}
  (i) \quad &0\leq|x|\leq H, \quad x\in\mathbb{Z},\\
  (ii) \quad &(\tau+x)(\tau-x)\equiv 0\mod 2\cdot 8^{m+1}, &\text{(II)}\\
  (iii) \quad &(H+x)(H-x)\equiv 0\mod K
\end{align*}

From~\autoref{lemma-2}, (I)\,(ii) satisfies if and only if (II)\,(iii) satisfies, since if $\tau-x
\equiv 0\mod 8^{m+1}$, then it satisfies (I)\,(ii); and if $\tau+x\equiv 0\mod 8^{m+1}$, then it
satisfies (i) and (iii).

Note how:

\begin{align*}
  (i) \quad &(\tau+x)(\tau-x)\equiv\tau^2-x^2\equiv 0\mod 2\cdot 8^{m+1}\\
  (ii) \quad &(H+x)(H-x)\equiv H^2-x^2\equiv 0\mod K
\end{align*}

So we can restrict condition (i) to only positive integers:

\begin{equation*}
  0\leq x_1\leq H, \quad x_1\in\mathbb{Z}
\end{equation*}

Also note that $\gcd(2\cdot 8^{m+1}, K)=1$, since $K$ is a product of odd primes (in fact greater
than 12) and $2\cdot 8^{m+1}$ a power of 2. Because of that, we can join both conditions (ii) and
(iii) from system (II).

\begin{equation*}
  \lambda_1 2\cdot 8^{m+1}(H^2-x_1^2)+\lambda_2 K(\tau^2-x_1^2)\equiv 0\mod 2\cdot 8^{m+1}\cdot K
\end{equation*}

Where $\lambda_1$ and $\lambda_2$ are parameters we can freely choose subject to the following
condition:

\begin{equation*}
  \gcd(\lambda_1,K)=\gcd(\lambda_2,2\cdot 8^{m+1})=1,\quad\lambda_1,\lambda_2\in\mathbb{Z}
\end{equation*}

We join these conditions into the new system (III):

\begin{align*}
  (i) \quad &0\leq x_1\leq H, \quad x_1\in\mathbb{Z},\\
  (ii) \quad &\lambda_1 2\cdot 8^{m+1}(H^2-x_1^2)+\lambda_2 K(\tau^2-x_1^2)\equiv 0\mod 2\cdot
    8^{m+1}\cdot K, &\text{(III)}\\
  (iii) \quad &\gcd(\lambda_1,K)=\gcd(\lambda_2,2\cdot 8^{m+1})=1,\quad\lambda_1,\lambda_2\in
    \mathbb{Z}
\end{align*}

In fact, choosing $\lambda_1$ and $\lambda_2$ provides proofs for two problems: the quadratic
Diophantine problem (which we will mention later) and for the QCP\@. Furthermore, conditions
(III)\,(i),\,(ii) are satisfiable for any $\lambda_1,\lambda_2$ obeying (III)\,(iii) or for no
$\lambda_1$ and $\lambda_2$.

We now finally provide a proof for~\autoref{qcp-thm}.

\begin{proof}[Proof of~\autoref{qcp-thm}]
  Choose $\lambda_1=\lambda_2=1$. It obviously satisfies (III)\,(iii). We rewrite (III)\,(ii) with
  these values:
  \begin{align*}
    &2\cdot 8^{m+1}(H^2-x_1^2)+K(\tau^2-x_1^2)\equiv 0\mod 2\cdot 8^{m+1}\cdot K\\
    &2\cdot 8^{m+1}H^2-2\cdot 8^{m+1}x_1^2+K\tau^2-x_1^2K\equiv 0\mod 2\cdot 8^{m+1}\cdot K\\
    &K\tau^2+2\cdot 8^{m+1}\cdot H^2\equiv(2\cdot 8^{m+1}+K)\cdot x_1^2\mod 2\cdot 8^{m+1}\cdot K
  \end{align*}
  Note how $\gcd(2\cdot 8^{m+1}+K,2\cdot 8^{m+1}\cdot K)=1$, and as such there exists an inverse
  ${(2\cdot 8^{m+1}+K)}^{-1}$. Multiplying the inverse on both sides, yields:

  \begin{equation*}
    {(2\cdot 8^{m+1}+K)}^{-1}\cdot(2\cdot 8^{m+1}+K)x_1^2\equiv{(2\cdot 8^{m+1}+K)}^{-1}(K\tau^2+
    2\cdot 8^{m+1}\cdot H^2)\mod 2\cdot 8^{m+1}\cdot K
  \end{equation*}

  Rewriting the expression above gives us:

  \begin{equation*}
    x_1^2\equiv {(2\cdot 8^{m+1}+K)}^{-1}(K\tau^2+2\cdot 8^{m+1}\cdot H^2)\mod 2\cdot 8^{m+1}\cdot K
  \end{equation*}

  Note how this is exactly the output of the algorithm for the QCP\@. Call $\alpha={(2\cdot
  8^{m+1}+K)}^{-1}(K\tau^2+2\cdot 8^{m+1}\cdot H^2)$ and $\beta=2\cdot 8^{m+1}\cdot K$. Rename
  $x_1$ as $x$. Note how we now have:

  \begin{equation*}
    x^2\equiv\alpha\mod\beta
  \end{equation*}

  Renaming $H$ as $\gamma$ gives us the restriction

  \begin{equation*}
    0\leq x^2\leq\gamma
  \end{equation*}

  where $\alpha,\beta,\gamma\in\omega$. But this is exactly the theorem. So the QCP is satisfiable
  if and only if system (III) is satisfiable. But (III) is satisfiable if and only if the
  propositional formula (i.e.\ 3-SAT) is satisfiable. From this we conclude that QCP is
  NP-complete.
\end{proof}

\section{Related problems}

Manders and Adleman proved in both~\cite{qcp1,qcp2} that not only the QCP is NP-complete, but also
the following:

\begin{theorem}\label{diophantine-thm}
  The (problem of accepting the) set of Diophantine equations (in a standard binary encoding) of
  the form
  \begin{equation*}
    \alpha x_1^2+\beta x_2-\gamma=0;\quad\alpha,\beta,\gamma\in\omega,
  \end{equation*}
  which have natural-number solutions $x_1,x_2$ is NP-complete.
\end{theorem}

The proof of this theorem starts exactly as the reduction we showed, with the only difference being
in the choice of $\lambda_1$ and $\lambda_2$. We leave it to the reader to prove this (take
$\lambda_1={(K+1)}^3$ and $\lambda_2=-1$).
